---
title: "c-rb-classification"
author: "vk"
date: "2026-04-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(terra)
library(rpart)
library(rpart.plot)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tree)
```

```{r reading in data and processing}
df <- read.csv('/Users/your path') ##put in your path to file here (data should be the training points with spectral sig)
df <- df%>%
  select(-".geo", -"system.index")
df$class <- factor(df$class,
                        levels = c(1,2,3,4,5,6),
                    labels = c("Fresh Sand", "Sand Mining", "Wet Sand", "Fallow", "Veg Sand", "Bedrock"))
```
```{r plotting spectral signature}
long <- pivot_longer(df, cols = -class, names_to = "band", values_to = "reflectance")
sand_summary <- long%>%
  group_by(class, band)%>%
  summarise(mean_reflectance = mean(reflectance), .groups = "drop")

ggplot(sand_summary, aes(x = band, y = mean_reflectance, color = class, group = class))+
  geom_line()+
  theme_minimal()+
  labs(title = "Spectral Signatures of Sand")+
  scale_color_discrete(labels = c("1" = "fresh alluv", "2" = "sandmine", "3" = "wet sand", "4" = "fallow", "5" = "veg sand", "6" = "rock"))

```
#########################################

model <- tree(class~ B11 + B12 + B2 + B3 + B4 + B8 + NDBI + NDMI + MNDWI + NDVI + BCI + contrast 
                      + correlation + dissimilarity + entropy + homogeneity + variance + asm +
                            NDSI, data = df)
plot(model)
text(model, pretty = 0, cex = 0.6)
cv <- cv.tree(model, FUN = prune.misclass)
cv

summary(model)
plot(cv$size,
     cv$dev,
     type = "b",
     xlab = "Tree Size",
     ylab = "CV Misclassification Error")

pruned <- prune.misclass(model, best = 6)

plot(pruned)
text(pruned, pretty = 0, cex = 0.7)
title("Classification and Regression Tree",
      col.main = "grey10",
      font.main = 2,
      cex.main = 0.9,
      adj = 0.6)
summary(pruned)
max_depth <- max(floor(log2(as.integer(rownames(pruned$frame)))))
print(max_depth)
print(pruned)

#Use the CART analysis to understand basis of classification
